---
title: "Support Vector Machines"
author: "Yifei Sun, Runze Cui"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
--- 

\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(mlbench)
library(ISLR)
library(caret)
library(tidymodels)
library(e1071)
library(kernlab)
library(ggrepel)
```




We use the Pima Indians Diabetes Database for illustration. The outcome is a binary variable `diabetes`. 

```{r}
data(PimaIndiansDiabetes2)
dat <- na.omit(PimaIndiansDiabetes2)
dat$diabetes <- factor(dat$diabetes, c("pos", "neg"))

set.seed(111111)
data_split <- initial_split(dat, prop = 0.75)

training_data <- training(data_split)
testing_data <- testing(data_split)
```



## Using `e1071`

Check https://cran.r-project.org/web/packages/e1071/vignettes/svmdoc.pdf for more details.

### Linear boundary

Most real data sets will not be fully separable by a linear boundary. Support vector classifiers with a tuning parameter `cost`, which quantifies the penalty associated with having an observation on the wrong side of the classification boundary, can be used to build a linear boundary.

```{r}
set.seed(1)
linear.tune <- tune.svm(diabetes ~ . , 
                        data = training_data, 
                        kernel = "linear", 
                        cost = exp(seq(-5,2, len = 50)),
                        scale = TRUE)
plot(linear.tune)
# summary(linear.tune)
linear.tune$best.parameters

best.linear <- linear.tune$best.model
summary(best.linear)

pred.linear <- predict(best.linear, newdata = testing_data)

confusionMatrix(data = pred.linear, 
                reference = testing_data$diabetes)

plot(best.linear, training_data, 
     glucose ~ mass,
     slice = list(pregnant = 5, triceps = 20,
                  insulin = 20, pressure = 75,
                  pedigree = 1, age = 50),
     grid = 100)
```

### Radial kernel

Support vector machines can construct classification boundaries that are nonlinear in shape. We use the radial kernel.

```{r}
set.seed(1)
radial.tune <- tune.svm(diabetes ~ . , 
                        data = training_data, 
                        kernel = "radial", 
                        cost = exp(seq(1, 7, len = 50)),
                        gamma = exp(seq(-10, -2,len = 20)))

plot(radial.tune, transform.y = log, transform.x = log, 
     color.palette = terrain.colors)
# summary(radial.tune)

radial.tune$best.parameters

best.radial <- radial.tune$best.model
summary(best.radial)

pred.radial <- predict(best.radial, newdata = testing_data)

confusionMatrix(data = pred.radial, 
                reference = testing_data$diabetes)

plot(best.radial, training_data, 
     glucose ~ mass,
     slice = list(pregnant = 5, triceps = 20,
                  insulin = 20, pressure = 75,
                  pedigree = 1, age = 50),
     grid = 100,
     symbolPalette = c("cyan","darkblue"),
     color.palette = heat.colors)
```   


## Using `caret`

```{r}
ctrl <- trainControl(method = "cv")

# kernlab
set.seed(1)
svml.fit <- train(diabetes ~ . , 
                  data = training_data, 
                  method = "svmLinear",
                  tuneGrid = data.frame(C = exp(seq(-5, 2, len = 50))),
                  trControl = ctrl)

plot(svml.fit, highlight = TRUE, xTrans = log)

# e1071
set.seed(1)
svml.fit2 <- train(diabetes ~ . , 
                   data = training_data, 
                   method = "svmLinear2",
                   tuneGrid = data.frame(cost = exp(seq(-5, 2, len = 50))),
                   trControl = ctrl)

plot(svml.fit2, highlight = TRUE, xTrans = log)
```


```{r, fig.width=15, fig.height=8}
svmr.grid <- expand.grid(C = exp(seq(1, 7, len = 50)),
                         sigma = exp(seq(-10, -2, len = 20)))

# tunes over both cost and sigma
set.seed(1)             
svmr.fit <- train(diabetes ~ . , data = training_data,
                  method = "svmRadialSigma",
                  tuneGrid = svmr.grid,
                  trControl = ctrl)

myCol <- rainbow(25)
myPar <- list(superpose.symbol = list(col = myCol),
              superpose.line = list(col = myCol))

plot(svmr.fit, highlight = TRUE, par.settings = myPar)

# tune over cost and uses a single value of sigma based on kernlab's sigest function
set.seed(1)             
svmr.fit2 <- train(diabetes ~ . , data = training_data,
                   method = "svmRadialCost",
                   tuneGrid = data.frame(C = exp(seq(-3, 3, len = 20))),
                   trControl = ctrl)

# Platt’s probabilistic outputs; use with caution
set.seed(1)             
svmr.fit3 <- train(diabetes ~ . , data = training_data,
                   method = "svmRadialCost",
                   tuneGrid = data.frame(C = exp(seq(-3, 3, len = 20))),
                   trControl = ctrl,
                   prob.model = TRUE) 
# predict(svmr.fit3, newdata = x_test, type = "prob")
```

```{r}
resamp <- resamples(list(svmr = svmr.fit, svmr2 = svmr.fit2,
                         svml = svml.fit, svml2 = svml.fit2))

summary(resamp)
bwplot(resamp)
```

We finally look at the test data performance.
```{r}
pred.svml <- predict(svml.fit, newdata = testing_data)
pred.svmr <- predict(svmr.fit, newdata = testing_data)

confusionMatrix(data = pred.svml, 
                reference = testing_data$diabetes)

confusionMatrix(data = pred.svmr, 
                reference = testing_data$diabetes)
```



## Using `tidymodels`

```{r}
set.seed(1)
cv_folds <- vfold_cv(training_data)

# Model specification
svm_linear_spec <- svm_linear(cost = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# svm_linear_spec %>% extract_parameter_dials("cost")

# Create a grid of tuning parameters
svm_linear_grid_set <- parameters(cost(range = c(-10, 5)))
svm_linear_grid <- grid_regular(svm_linear_grid_set, levels = 50)

# Create a workflow
svm_linear_workflow <- workflow() %>%
  add_model(svm_linear_spec) %>% 
  add_formula(diabetes ~ .)

# Tune the model
svm_linear_tune <- tune_grid(
  svm_linear_workflow,
  resamples = cv_folds,
  grid = svm_linear_grid)

# Plot the results
autoplot(svm_linear_tune, metric = "accuracy")

svm_linear_best <- select_best(svm_linear_tune, metric = "accuracy")

final_svm_linear_spec <- svm_linear_spec %>% 
  update(cost = svm_linear_best$cost)

svm_linear_fit <- fit(final_svm_linear_spec, formula = diabetes ~ ., data = training_data)

svm_linear_model <- extract_fit_engine(svm_linear_fit)

head(predict(svm_linear_fit, new_data = testing_data))
```

```{r}
# Model specification
svmrbf_spec <- svm_rbf(cost = tune(), rbf_sigma = tune()) %>%
  set_engine("kernlab") %>%
  set_mode("classification")

# svmrbf_spec %>% extract_parameter_dials("cost")
# svmrbf_spec %>% extract_parameter_dials("rbf_sigma")

# Create a grid of tuning parameters
svmrbf_grid_set <- parameters(cost(range = c(1, 10), trans = log_trans()), 
                              rbf_sigma(range = c(-10, 0), trans = log_trans()))
svmrbf_grid <- grid_regular(svmrbf_grid_set, levels = c(10, 20))

# Create a workflow
svmrbf_workflow <- workflow() %>%
  add_model(svmrbf_spec) %>% 
  add_formula(diabetes ~ .)

# Tune the model
svmrbf_tune <- tune_grid(
  svmrbf_workflow,
  resamples = cv_folds,
  grid = svmrbf_grid)

# Plot the results
autoplot(svmrbf_tune, metric = "accuracy")

svmrbf_best <- select_best(svmrbf_tune, metric = "accuracy")

final_svmrbf_spec <- svmrbf_spec %>% 
  update(cost = svmrbf_best$cost, rbf_sigma = svmrbf_best$rbf_sigma)

svmrbf_fit <- fit(final_svmrbf_spec, formula = diabetes ~ ., data = training_data)

svmrbf_model <- extract_fit_engine(svmrbf_fit)

head(predict(svmrbf_fit, new_data = testing_data))
```

```{r}
model_compare <- workflow_set(preproc = list(diabetes ~ .),
                              models = list(svm_linear = final_svm_linear_spec, 
                                            svm_rbf = final_svmrbf_spec)) %>% 
  workflow_map(resamples = cv_folds)

model_compare %>% collect_metrics() %>% filter(.metric == "accuracy")
```

