---
title: "Ridge Regression"
author: "Yifei Sun, Runze Cui"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '2'
header-includes:
- \usepackage{fancyhdr}
- \usepackage{lipsum}
- \pagestyle{fancy}
- \fancyhead[R]{\thepage}
- \fancypagestyle{plain}{\pagestyle{fancy}}
editor_options: 
  chunk_output_type: console
--- 

\newpage


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = T, message = FALSE, results='hide', warning=FALSE}
library(ISLR)
library(glmnet)
library(caret)
library(tidymodels)
library(corrplot)
library(ggplot2)
library(plotmo)
library(ggrepel)
```

Predict a baseball player’s salary on the basis of various statistics associated with performance in the previous year. Use `?Hitters` for more details.

```{r}
data(Hitters)
Hitters <- na.omit(Hitters)

set.seed(2222)
data_split <- initial_split(Hitters, prop = 0.8)

# Extract the training and test data
training_data <- training(data_split)
testing_data <- testing(data_split)
```

# Using `glmnet`

## Ridge regression

```{r}
# matrix of predictors (glmnet uses input matrix)
x <- model.matrix(Salary ~ ., training_data)[,-1]
# vector of response
y <- training_data[, "Salary"]

corrplot(cor(x), method = "circle", type = "full")
```

`alpha` is the elastic net mixing parameter. `alpha=1` is the lasso penalty, and `alpha=0` the ridge penalty. `glmnet()` function standardizes the independent variables by default (The coefficients are always returned on the original scale). 

```{r}
# fit the ridge regression (alpha = 0) with a sequence of lambdas
ridge.mod <- glmnet(x = x, y = y, 
                    # standardize = TRUE,
                    alpha = 0, 
                    lambda = exp(seq(10, -5, length = 100)))
```

`coef(ridge.mod)` gives the coefficient matrix. Each column is the fit corresponding to one lambda value.

```{r}
mat.coef <- coef(ridge.mod)
dim(mat.coef)
```


### Trace plot

```{r}
# plot(ridge.mod, xvar = "lambda", label = TRUE)
plot_glmnet(ridge.mod, xvar = "rlambda", label = 19)
```

### Cross-validation

We use cross-validation to determine the optimal value of `lambda`. The two vertical lines are the for minimal MSE and 1SE rule. The 1SE rule gives the most regularized model such that error is within one standard error of the minimum.

```{r}
## ridge alpha = 0
set.seed(2)
cv.ridge <- cv.glmnet(x, y, 
                      alpha = 0, 
                      lambda = exp(seq(10, -5, length = 100)))
# set.seed(2)
# cv.ridge <- cv.glmnet(x, y, alpha = 0, nlambda = 200)

plot(cv.ridge)
abline(h = (cv.ridge$cvm + cv.ridge$cvsd)[which.min(cv.ridge$cvm)], col = 4, lwd = 2)
## up number = nonzero parameter number

# min CV MSE
cv.ridge$lambda.min
# the 1SE rule
cv.ridge$lambda.1se
```


### Coefficients of the final model

Get the coefficients of the optimal model. `s` is value of the penalty parameter `lambda` at which predictions are required.

```{r}
# extract coefficients
predict(cv.ridge, s = cv.ridge$lambda.min, type = "coefficients") 

# make prediction
head(predict(cv.ridge, newx = model.matrix(Salary ~ ., testing_data)[,-1], 
             s = "lambda.min", type = "response")) 

# predict(cv.ridge, s = "lambda.min", type = "coefficients") 
# predict(cv.ridge, s = "lambda.1se", type = "coefficients") 
# predict(ridge.mod, s = cv.ridge$lambda.min, type = "coefficients")
```

## Lasso 

The syntax is along the same line as ridge regression. Now we use `alpha = 1`.

```{r}
## lasso alpha = 1
cv.lasso <- cv.glmnet(x, y, 
                      alpha = 1, 
                      lambda = exp(seq(6, -5, length = 100)))

cv.lasso$lambda.min
```

```{r}
plot(cv.lasso)
## u should avoid the case where the lambda that gives you the smallest cross validation error is on the boundary of your grid.
## So if that is the case, it means that you should expand the grid,
```

```{r}
# cv.lasso$glmnet.fit is a fitted glmnet object using the full training data
# plot(cv.lasso$glmnet.fit, xvar = "lambda", label=TRUE)
plot_glmnet(cv.lasso$glmnet.fit)
```


```{r}
predict(cv.lasso, s = "lambda.min", type = "coefficients")

head(predict(cv.lasso, newx = model.matrix(Salary ~ ., testing_data)[,-1], 
             s = "lambda.min", type = "response"))
```


# Using `caret`

## Ridge regression

```{r}
ctrl1 <- trainControl(method = "cv", number = 10)

set.seed(2)
ridge.fit <- train(Salary ~ . ,
                   data = training_data,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 0,
                                          lambda = exp(seq(6, 0, length=100))),
                   ## expand dot grid function is used to create a data frame that contains all the combination of tune parameters that you that you specify
                   trControl = ctrl1)

plot(ridge.fit, xTrans = log)


ridge.fit$bestTune

# coefficients in the final model
coef(ridge.fit$finalModel, s = ridge.fit$bestTune$lambda)
```

## Lasso

```{r}
set.seed(2)
lasso.fit <- train(Salary ~ .,
                   data = training_data,
                   method = "glmnet",
                   tuneGrid = expand.grid(alpha = 1, 
                                          lambda = exp(seq(6, 0, length = 100))),
                   trControl = ctrl1)
plot(lasso.fit, xTrans = log)

lasso.fit$bestTune

# coefficients in the final model
coef(lasso.fit$finalModel, lasso.fit$bestTune$lambda)
```

## Elastic net

```{r}
set.seed(2)
enet.fit <- train(Salary ~ .,
                  data = training_data,
                  method = "glmnet",
                  tuneGrid = expand.grid(alpha = seq(0, 1, length = 21), 
                                         lambda = exp(seq(6, 0, length = 100))),
                  trControl = ctrl1)
enet.fit$bestTune

myCol <- rainbow(25)
myPar <- list(superpose.symbol = list(col = myCol),
              superpose.line = list(col = myCol))

plot(enet.fit, par.settings = myPar)

# coefficients in the final model
coef(enet.fit$finalModel, enet.fit$bestTune$lambda)
```

## Comparing different models

```{r, fig.width=5}
set.seed(2)
lm.fit <- train(Salary ~ .,
                data = training_data,
                method = "lm",
                trControl = ctrl1)

resamp <- resamples(list(enet = enet.fit, lasso = lasso.fit, ridge = ridge.fit, lm = lm.fit))

summary(resamp)

parallelplot(resamp, metric = "RMSE")
# bwplot(resamp, metric = "RMSE")
```

## Prediction

```{r}
enet.pred <- predict(enet.fit, newdata = testing_data)
# test error
mean((enet.pred - testing_data[, "Salary"])^2)
```


# Using `tidymodels`

## Ridge regression

```{r}
set.seed(2)
cv_folds <- vfold_cv(training_data, v = 10) 

# Model specification for ridge regression
ridge_spec <- linear_reg(penalty = tune(), mixture = 0) %>% # mixture = 0 for ridge regression
  set_engine("glmnet") %>% 
  set_mode("regression")

# ridge_spec %>% extract_parameter_dials("penalty")

# Grid of tuning Parameters
ridge_grid_set <- parameters(penalty(range = c(-2, 5), trans = log_trans()))
ridge_grid <- grid_regular(ridge_grid_set, levels = 100)

# Set up the workflow
ridge_workflow <- workflow() %>%
  add_model(ridge_spec) %>%
  add_formula(Salary ~ .)

# Tune the model
ridge_tune <- tune_grid(
  ridge_workflow,
  resamples = cv_folds,
  grid = ridge_grid,
  control = control_resamples(extract = extract_fit_parsnip, save_pred = TRUE)
)

# CV plot
autoplot(ridge_tune, metric = "rmse") 

# Select tuning parameters based on 1SE rule
ridge_1SE <- select_by_one_std_err(ridge_tune, metric = "rmse", desc(penalty)) 

# !!! see "why_is_ridge_CV_curve_flat.R"
ridge_best <- select_best(ridge_tune, metric = "rmse") 
cv_rmse <- ridge_tune %>% collect_metrics() %>% filter(.metric == "rmse") 
cv_rmse_mean <- cv_rmse$mean
which(cv_rmse_mean == min(cv_rmse_mean))

# Update the model with the best lambda
final_ridge_spec <- ridge_spec %>% 
  update(penalty = 1.3)

# Fit your final model to the train data
ridge_fit <- fit(final_ridge_spec, formula = Salary ~ ., data = training_data)

# Get coefficients
ridge_model <- extract_fit_engine(ridge_fit)
coef(ridge_model, s = ridge_1SE$penalty)
```


## Lasso

```{r}
lasso_spec <- linear_reg(penalty = tune(), mixture = 1) %>% # mixture = 1 for lasso regression
  set_engine("glmnet") %>% 
  set_mode("regression")

# lasso_spec %>% extract_parameter_dials("penalty")

lasso_grid_set <- parameters(penalty(range = c(-3, 5), trans = log_trans()))
lasso_grid <- grid_regular(lasso_grid_set, levels = 100)

lasso_workflow <- workflow() %>%
  add_model(lasso_spec) %>%
  add_formula(Salary ~ .)

lasso_tune <- tune_grid(
  lasso_workflow,
  resamples = cv_folds,
  grid = lasso_grid
)

autoplot(lasso_tune, metric = "rmse")

lasso_best <- select_best(lasso_tune, metric = "rmse") 

final_lasso_spec <- lasso_spec %>% 
  update(penalty = lasso_best$penalty)

lasso_fit <- fit(final_lasso_spec, formula = Salary ~ ., data = training_data)

lasso_model <- extract_fit_engine(lasso_fit)
coef(lasso_model, s = lasso_best$penalty)
```

## Elastic net

```{r}
enet_spec <- linear_reg(penalty = tune(), mixture = tune()) %>% 
  set_engine("glmnet") %>% 
  set_mode("regression")

# enet_spec %>% extract_parameter_dials("mixture")
# enet_spec %>% extract_parameter_dials("mixture")

enet_grid_set <- parameters(penalty(range = c(-3, 5), trans = log_trans()),
                            mixture(range = c(0, 1)))
enet_grid <- grid_regular(enet_grid_set, levels = c(100, 21))



enet_workflow <- workflow() %>%
  add_model(enet_spec) %>%
  add_formula(Salary ~ .)

enet_tune <- tune_grid(
  enet_workflow,
  resamples = cv_folds,
  grid = enet_grid
)

autoplot(enet_tune, metric = "rmse") + 
  theme(legend.position = "top") +
  labs(color = "Mixing Percentage\n(Alpha Values)") 

enet_best <- select_best(enet_tune, metric = "rmse") 

final_enet_spec <- enet_spec %>% 
  update(penalty = enet_best$penalty, mixture = enet_best$mixture)

enet_fit <- fit(final_enet_spec, formula = Salary ~ ., data = training_data)

# Get coefficients
enet_model <- extract_fit_engine(enet_fit)
coef(enet_model, s = enet_best$penalty)
```

## Comparing different models

```{r}
lm_spec <- linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")

model_compare <- workflow_set(preproc = list(Salary ~ .),
                              models = list(lm = lm_spec, 
                                            lasso = final_lasso_spec,
                                            ridge = final_ridge_spec,
                                            enet = final_enet_spec)) %>% 
  workflow_map(resamples = cv_folds) 

autoplot(model_compare, metric = "rmse") +
  geom_text_repel(aes(label = wflow_id, color = wflow_id), 
                  nudge_x = 1/8, nudge_y = 1/100) +
  theme(legend.position = "none")
```

## Prediction

```{r}
enet_pred <- predict(enet_fit, new_data = testing_data)

# Calculate test RMSE
sqrt(mean((enet_pred[[1]] - testing_data$Salary)^2))
```


