---
title: "8106hw5-1"
author: "Ze Li"
output: pdf_document
---

```{r library, echo = T, message = FALSE, results='hide', warning=FALSE}
library(ISLR)
library(caret)
library(glmnet)
library(tidymodels)
library(mlbench)
library(pROC)
library(pdp)
library(vip)
library(AppliedPredictiveModeling)
library(MASS)
library(ggplot2)
```

## Problem 1

```{r data import}
auto = read.csv("/Users/zeze/Library/Mobile Documents/com~apple~CloudDocs/2024/24S BIST P8106 DS II/hw5/auto.csv")
head(auto)
data_split <- initial_split(auto, prop = 0.7)
train <- training(data_split) 
test <- testing(data_split)
x_test <- model.matrix(mpg_cat ~ ., test)[, -1]
head(train)
```

### (a) Fit a support vector classifier to the training data. What are the training and test error rates?

```{r svc fit}
ctrl <- trainControl(method = "cv")
# kernlab
set.seed(1)
svml.fit <- train(mpg_cat ~ . , data = train,
                  method = "svmLinear",
                  tuneGrid = data.frame(C = exp(seq(-5, 2, len = 50))),
                  trControl = ctrl)
plot(svml.fit, highlight = TRUE, xTrans = log)
# e1071
set.seed(1)
svml.fit2 <- train(mpg_cat ~ . , data = train,
                   method = "svmLinear2",
                   tuneGrid = data.frame(cost = exp(seq(-5, 2, len = 50))),
                   trControl = ctrl)
plot(svml.fit2, highlight = TRUE, xTrans = log)
```

```{r svc error rate}
resamp <- resamples(list(svml = svml.fit, svml2 = svml.fit2))
summary(resamp)
pred.svml <- predict(svml.fit, newdata = test)
# Ensure both are factors and have the same levels in the same order
train$mpg_cat <- factor(train$mpg_cat, levels = c("high", "low"))
test$mpg_cat <- factor(test$mpg_cat, levels = c("high", "low"))
pred.svml <- factor(pred.svml, levels = c("high", "low"))
# train error rate
confusionMatrix(data = pred.svml, reference = train$mpg_cat)
# test error rate
confusionMatrix(data = pred.svml, reference = test$mpg_cat)
```

The accurarcy is 90.68%, so the error rate is 9.32%.

The best linear fit visualization is 

```{r}
plot(best.linear, data = train,
     weight ~ horsepower,
     slice = list(displacement = 8, cylinders = 8,
                  acceleration = 18, year = 72,
                  origin = 2),
     grid = 50)
```

### (b) Fit a support vector machine with a radial kernel to the training data. What are the training and test error rates?

```{r svm fit}
svmr.grid <- expand.grid(C = exp(seq(1, 7, len = 50)),
                         sigma = exp(seq(-10, -2, len = 20)))
# tunes over both cost and sigma
set.seed(1)
svmr.fit <- train(mpg_cat ~ . , data = train,
                  method = "svmRadialSigma",
                  tuneGrid = svmr.grid,
                  trControl = ctrl)
myCol <- rainbow(25)
myPar <- list(superpose.symbol = list(col = myCol),
              superpose.line = list(col = myCol))
plot(svmr.fit, highlight = TRUE, par.settings = myPar)
```

```{r svm fit2}
# tune over cost and uses a single value of sigma based on kernlab's sigest function
set.seed(1)
svmr.fit2 <- train(mpg_cat ~ . , data = train,
                   method = "svmRadialCost",
                   tuneGrid = data.frame(C = exp(seq(-3, 3, len = 20))),
                   trControl = ctrl)
# Platt’s probabilistic outputs; use with caution
set.seed(1)
svmr.fit3 <- train(mpg_cat ~ . , data = train,
                   method = "svmRadialCost",
                   tuneGrid = data.frame(C = exp(seq(-3, 3, len = 20))),
                   trControl = ctrl,
                   prob.model = TRUE)
predict(svmr.fit3, newdata = x_test, type = "prob")
resamp2 <- resamples(list(svmr = svmr.fit, svmr2 = svmr.fit2))
summary(resamp2)
```

```{r svm error rate}
pred.svmr <- predict(svmr.fit, newdata = test)
pred.svmr <- factor(pred.svmr, levels = c("high", "low"))
# train error rate
confusionMatrix(data = pred.svmr, reference = train$mpg_cat)
# test error rate
confusionMatrix(data = pred.svmr, reference = test$mpg_cat)
```

The accurarcy is 91.53%, so the error rate is 8.48%.

The best radial visualization is

```{r}
plot(best.radial, train,
     displacement ~ weight,
     slice = list(cylinders = 8, horsepower = 100,
                  acceleration = 18, year = 72,
                  origin = 2),
     grid = 100,
     symbolPalette = c("cyan","darkblue"),
     color.palette = heat.colors)
```


## Problem 2

```{r data2}
data("USArrests")
USArrests = USArrests %>%
  as_tibble()
head(USArrests)
```

### (a) Using hierarchical clustering with complete linkage and Euclidean distance, cluster the states. Cut the dendrogram at a height that results in three distinct clusters. Which states belong to which clusters?

```{r}

```


### (b) Hierarchically cluster the states using complete linkage and Euclidean distance, after scaling the variables to have standard deviation one.

### (c) Does scaling the variables change the clustering results? Why? In your opinion, should the variables be scaled before the inter-observation dissimilarities are computed?
